---
title: "NWSL Preliminary Data Viewing"
author: "Hannah Mandell"
date: "1/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
#install.packages("devtools")
library(devtools)
devtools::install_github("adror1/nwslR")

library(ggplot2)
library(lubridate)
library(dplyr)
library(tidyr)
library(broom)
library(praise)
library(nwslR)
library(plotly)
```

```{r loading data}
player_season_stats <- nwslR::fieldplayer_overall_season_stats
team_data <- nwslR::team_stats_season
draft_data <- nwslR::draftpicks
adv_player_stats <- nwslR::adv_player_stats	
adv_team_stats <- nwslR::adv_team_stats	
awards <- nwslR::award
franchise <- nwslR::franchise
players <- nwslR::player

```

# Goal Efficiency

```{r pk analysis}
nwsl_goals_per_min <- player_season_stats %>%
  group_by(player) %>%
  summarise(goals_per_min = gls/min)

nwsl_goals_per_min_career <- nwsl_goals_per_min %>%
  group_by(player) %>%
  summarise(goals_per_min = mean(goals_per_min))
```


```{r joining the datasets}
player_season_stats <- inner_join(basic_player_data, player_season_stats, by = "person_id", copy = FALSE) %>%
  select(-nation.y, -pos.y, -name_other)
```

```{r a look into carli lloyd}
players %>%
  filter(player == "Carli Lloyd")


player_season_stats %>%
  filter(player == "Carli Lloyd")

```

```{r}
player_season_stats %>%
  filter(player == c("Carli Lloyd")) %>%
  ggplot(aes(x = season, y = gls)) + 
  geom_point(color = "red")

```


```{r}
player_season_stats <- player_season_stats %>%
  rename(pos = pos.x, nation = nation.x)
```

```{r do selfish players score more?}
p1 <- player_season_stats %>%
  #group_by(team_id) %>%
  ggplot(aes(x = gls, y = ast)) +
  geom_point(color = "blue") + 
  geom_smooth(color = "red")

ggplotly(p1, )

```
There appears to be a weak positive correlation between goals and assists per season. The correlation grows weaker as the player scores more goals, meaning that they assist less. Let's look at this by position:


```{r}
player_season_stats %>%
  group_by(pos) %>%
  summarise(count = n())
```
```{r}
ggplotly(
  player_season_stats %>%
  group_by(pos) %>%
  #mutate(gls = median(gls), ast = median(ast)) %>%#
  ggplot(aes(x =  pos, y = (ast/gls))) +
  geom_boxplot(outlier.shape = NA)
  )


```
```{r looking team-by-team}



maxgoals <- team_data %>%
  group_by(team_id) %>%
  group_by(season) %>%
  summarize(max(goals)) %>%
  ggplot(x = year, y = )



maxgoals

p2 <- 
  ggplot(maxgoals, aes(x = year, y = max(goals))
         
ggplotly(p2)

```

```{r max goals in any season}
maxgoals <- team_data %>%
  group_by(team_id) %>%
  summarize(max(goals))
  
maxgoals
```

```{r all goals}

all_goals <- team_data %>%
  group_by(season) %>%
  ggplot(aes(x = season, y = goals)) +
  geom_point(color = "red") +
  geom_smooth(color = "black")

ggplotly(all_goals)

```

There does not appear to be a significant trend in overal goal scorring from 2016 to 2019. 2016, however, is significantly lower than 2017, 2018, and 2019, which are all very similar in median values. 

# The Draft

Let's explore some of the rounds & popular feeder university programs. 

```{r first round feeders}

top_feeders <- draft_data %>%
  filter(round == 1) %>%
  count(previous_team, sort = TRUE)

top_feeders
```
UCLA is the top provider of 1st round NWSL draft picks since 2013, closely followed by FSU and Stanford. Let's explore the draft as a whole:

```{r all rounds feeders}

draft_colleges <- draft_data %>%
    count(previous_team, sort = TRUE)

draft_colleges

```
UCLA is again the top contender, but Penn State situates itself in second place's position, providing 12 NWSL draft picks since 2013. 

# Passing Accuracy

```{r creating a full name variable}

adv_player_stats$player_name = paste(adv_player_stats$first_name, adv_player_stats$last_name)
  
head(adv_player_stats$player_name)
```

```{r calculate passing accuracy for each game}
adv_player_stats$passing_acc = adv_player_stats$accurate_pass / adv_player_stats$total_pass

head(adv_player_stats$passing_acc)
```


```{r passing accuracy!}

passing_accuracy <- adv_player_stats %>%
  group_by(player_name) %>%
  summarize(avg_passing_acc = mean(passing_acc))

passing_accuracy <- passing_accuracy %>%
  arrange(desc(avg_passing_acc))

passing_accuracy
```
The NWSL players with the statistically highest passing accuracy are: 1) Lisa De Vanna, 2) Samantha Staab, and 3) Hope Solo. Let's check them out:

```{r}

player_name_and_id <- data_frame(players$person_id, players$player)

player_name_and_id

player_name_and_id <- player_name_and_id %>%
  rename(person_id = "players$person_id", player = "players$player")

```

```{r putting names into other dataset}
player_season_stats <- full_join(player_name_and_id, player_season_stats, by = "person_id")
```


```{r}
player_season_stats %>%
  filter(player == "Lisa De Vanna")

adv_player_stats %>%
  filter(player_name == "Lisa De Vanna")

adv_player_stats %>%
  filter(player_name == "Samantha Staab")

adv_player_stats %>%
  filter(player_name == "Hope Solo")


```

After some further investigation, we find that De Vanna has only played in 3 games, rendering her passing accuracy score less than completely picturesque of her footballer skills. Although her accuracy for those 3 games in 2016 is certanly notable, it is not entirely fair to compare them again Staab's 24, or even Solo's 8. I might place a filter on number of entries so we can compare passing accuracy across players that have played many games and therfore have had their passing tested under many circumstances. 

